---
title: 'Homework #2'
author: "Rachel Lee"
date: "10/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load libraries
```{r}
library(tidyverse)
library(tidyr)
library(readxl)
```


#Problem 1

```{r p1_import_clean}
trashwheel = read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                        sheet = 1, 
                        range = "A2:N408") %>% 
  janitor::clean_names() %>%
  select(-dumpster) %>%
  mutate(sports_balls = as.integer(round(sports_balls))) %>% 
  rename("weight" = weight_tons, "volume" = volume_cubic_yards) #rename columns to cleaner variable names
```

weight_tons and volume_cubic_yards were renamed to weight and volume respectively. It is not necessary to include units in variable names. 
Import and clean precipitation data from 2017 and 2018 


## Precipitation Dataset
The following code trunk reads and cleans precipitation data for 2017 and 2018. For each, omit rows without precipitation data and add a variable year. Next, combine precipitation datasets and convert month to a character variable. 

```{r p1_dataimport_prcp}
prcp_2017 = 
    read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = '2017 Precipitation', range = "A2:B14") %>%
    janitor::clean_names() %>% 
    drop_na(total) %>% #delete rows that do not contain precipitation data
    mutate(year = 2017)
    
prcp_2018 = 
    read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = '2018 Precipitation', range = "A2:B14") %>%
    janitor::clean_names() %>% 
    drop_na(total) %>%  #delete rows that do not contain precipitation data
    mutate(year = 2018)

# Combine datasets for years 2017 and 2018
prcp = bind_rows(prcp_2017, prcp_2018) %>% 
    mutate(month = month.name[month]) %>% 
    select(year, month, total)

# Confirm the combined dataset
head(prcp)
tail(prcp)
```
