---
title: 'Homework #2'
author: "Rachel Lee"
date: "10/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load libraries
```{r}
library(tidyverse)
library(tidyr)
library(readxl)
```


# Problem 1

```{r p1_import_clean}
trashwheel = read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                        sheet = 1, 
                        range = "A2:N408") %>% 
  janitor::clean_names() %>%
  select(-dumpster) %>%
  mutate(sports_balls = as.integer(round(sports_balls))) %>% 
  rename("weight" = weight_tons, "volume" = volume_cubic_yards) #rename columns to cleaner variable names
```

weight_tons and volume_cubic_yards were renamed to weight and volume respectively. It is not necessary to include units in variable names. 
Import and clean precipitation data from 2017 and 2018 


## Precipitation Dataset
The following code trunk reads and cleans precipitation data for 2017 and 2018. For each, omit rows without precipitation data and add a variable year. Next, combine precipitation datasets and convert month to a character variable. 

```{r p1_dataimport_prcp}
prcp_2017 = 
    read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = '2017 Precipitation', range = "A2:B14") %>%
    janitor::clean_names() %>% 
    drop_na(total) %>% #delete rows that do not contain precipitation data
    mutate(year = 2017)
    
prcp_2018 = 
    read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = '2018 Precipitation', range = "A2:B14") %>%
    janitor::clean_names() %>% 
    drop_na(total) %>%  #delete rows that do not contain precipitation data
    mutate(year = 2018)

# Combine datasets for years 2017 and 2018
prcp = bind_rows(prcp_2017, prcp_2018) %>% 
    mutate(month = month.name[month]) %>% 
    select(year, month, total)

# Confirm the combined dataset
head(prcp)
tail(prcp)
```
# Interpretation
In the Mr.Trash Wheel dataset, there are `r nrow(trashwheel)` observations and `r ncol(trashwheel)` variables. Key variables of the Mr. Trash Wheel dataset are  `colnames(trashwheel)`. In the precipitation datatset, there are `r nrow(prcp)` observations and `r ncol(prcp)` variables. Key variable of the precipitation dataset is `total` which is the precipitation volume for each month. `r sum(filter(prcp_2017, year == 2017)$total)` was the total precipitation in 2017. Median number of sports balls in 2018 was `r median(filter(trashwheel, year == 2018)$sports_balls)`.

# Problem 2

```{r p2_import_clean}

# cleaning polls dataset

polls = read_csv(file = "./pols-month.csv")%>%
  separate(mon, c("year", "month", "day"), 
           convert = TRUE)%>%
  mutate (month = month.name[month])%>%
  mutate(prez_dem = recode (prez_dem, '1' = "dem")) %>% 
  mutate(prez_gop = recode (prez_gop, '1' = "gop", '2' = "gop")) %>% 
  pivot_longer(
    cols = c("prez_dem", "prez_gop"), 
    values_to = "president", 
    values_drop_na = TRUE
  ) %>% 
  select(-day, -name)

# clean the snp dataset

snp = read_csv(file = "./snp.csv")%>%
    separate(date, c("year", "month", "day"), 
             convert = TRUE)%>%
   mutate (month = month.name[month])%>%
  select(year, month, -day, everything())

# clean and transform the unemployment data

unemployment = read_csv("./unemployment.csv") %>%
  pivot_longer(Jan:Dec, 
    names_to = "month",
    values_to = "rate"
  ) %>% 
  mutate(month = match(month, month.abb)) %>%
  mutate(month = month.name[month]) %>%
  mutate(Year = as.integer(Year)) %>%
  janitor::clean_names()


#combine data sets -> create one dataset with polls, snp and unemployent rate

Final = polls%>%
  left_join( snp,
             by = c("year", "month"))%>%
  left_join( unemployment,
             by = c("year", "month"))
```

# Interpretation
In the polls data set, there are `r nrow(polls)` observations. In the snp data set, there are `r nrow(snp)` observations. In the unemployment data set, there are `r nrow(unemployment)` observations. In the Final combined dataset, there are `r ncol(Final)` observations. Years range from `r min(pull(polls, year))` to `r max(pull(polls, year))`. 


# Problem 3

```{r}
popular_names = read_csv(file="./Popular_Baby_Names.csv")%>%
  janitor::clean_names()%>%
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "BLACK NON HISP" = "BLACK NON HISPANIC", "WHITE NON HISP" = "WHITE NON HISPANIC"))%>%
  mutate(childs_first_name = str_to_lower(childs_first_name))%>%
  distinct()
popular_names %>%
  filter(childs_first_name == "olivia")%>%
  select(-count)%>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank")
popular_names %>%
  filter(gender == "male")%>%
  filter(rank == 1)%>%
  select(-count)%>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  )
popular_names %>%
  filter(gender == "MALE",
         ethnicity == "WHITE NON HISPANIC",
         year_of_birth == 2016)%>%
  ggplot(aes(x=rank, y=count)) + 
  geom_point(colour = "purple") +
  labs(
    title = "Popular White-non Hispanic Male Names and Ranks ",
    x = "Popularity Ranking",
    y = "Count"
  )
  
```

